---
title: 'Psych 251 Comprehension Pilot A: Gick (1983)'
author: "Greses Perez"
date: "October 28th, 2018"
output:
  html_document:
    toc: yes
    toc_float: yes
  pdf_document:
    toc: yes
  word_document:
    toc: yes
---

This is pilot A is a pilot of my study with non-naive participants. In this document, you will find: 

1. The "data" collected from me and my friends to guarantee that the data is logging correctly. 
2. The feedback on the paradigm that I got by running it several times. 
3. The code for my planned analyses
4. The confirmation that I can run the code on my data.
5. The rendered replication report.
6. The link to my paradigm. 
7. The limited data that I collected analyzed via the confirmatory analyses.
8. The different conditions included in the paradigm. In the case of this study, there are 6 different conditions. 


First, let's look at the "data" collected from me and my friends to guarantee that the data is logging correctly.

```{r}
library(readr)
d <- read_csv("~/Desktop/PhD Stanford /PSYCH-251 Experimental Methods/Final Project /PilotA/PilotA.csv")
View(d)
cols_condense(d)
spec(d)
```

```{r}
D <- transform(d, A1 = as.numeric (A1),
               A2 = as.numeric (A2),
               A3 = as.numeric (A3),
               A4 = as.numeric (A4),
               A5 = as.numeric (A5),
               A6 = as.numeric (A6))
sapply(D, mode)
```

```{r}
sapply(D, class)
```



```{r}
colnames(d)
```


Take a look at the data structure:

```{r}
head(d)
```

```{r}
rd <- data.frame(D = c(0,1),
                cond = c(A,B,C,D,E,F),
                cond = c(1,2,3,4,5,6))
```

#### Here I am trying to organize the data and calculate the frequencies, but I have not been able to figure out the code to do so... I have come up with a way to do it modifying the original data file directly, but then I lose the purpose of this exercise which it is to try to do it in R. I need more time to figure this out. 


[link to Conditions included in Qualtrics] ("https://github.com/greses/gick1983.git/gick1983/DataPilotA/Conditions_Qualtrics.doc")
